Distributed Web Mining of Ethereum

### 2018-04-10 paper/paper.tex
final CSCML paper

% remove pseudocode.
% Have a section of ethereum and define what block header and hash. Also define target etc 
% dEFINE epoch early on. 
% called Cache in ethereum terminology. Mention that the size of the cache varies right 
% define lightweight and full node earlier.
% Say each DAG consists of nodes and slices. Explain what 
% Ads: talk about how our approach can be applied to

@@

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{float}
\usepackage{graphicx}
\usepackage{balance}
\usepackage{color}
\usepackage{makecell}
\usepackage{amsmath}
\usepackage{algorithmic}
\usepackage{amssymb}
\usepackage{caption}
\newcommand{\trishita}[1]{}%{{\color{magenta}\bfseries[Trishita: #1]}}
\newenvironment{claimproof}[1]{\par\noindent\underline{Proof:}\space#1}{\hfill $\blacksquare$}

% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}
\begin{document}
%
\title{Distributed Web Mining of Ethereum
% \thanks{Supported by organization x.}
}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{
Trishita Tiwari%\orcidID{0000-1111-2222-3333} 
\and
David Starobinski%\orcidID{1111-2222-3333-4444} 
\and
Ari Trachtenberg%\orcidID{2222--3333-4444-5555}
}
%
\authorrunning{T. Tiwari et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Boston University, Department of Electrical and Computer Engineering, Boston, MA 02215, USA \\
\email{\{trtiwari,staro,trachten\}@bu.edu}}
%
\maketitle             
%
\begin{abstract}
We consider the problem of mining crytocurrencies by harnessing the inherent distribution capabilities of the World Wide Web. More specifically, we consider
the mining of the Ethereum cryptocurrency by apportioning elements of the corresponding Proof-of-Work calculations among a number of web clients. These calculations are handled through individualized code that runs on the client browsers, and is thereafter collated at a web server to complete a mining operation. WebEth is based on a lazy evaluation technique designed to function within the expected limitations of the clients, including bounds on memory, computation and communication bandwidth to the server. We provide proofs-of-concept based in JavaScript and WebAssembly, with the latter reaching hash rates up to roughly 40 kiloHashes per second, which is only 30\% slower than the corresponding native C++-based miner.
Finally, we explore several applications for which this hash rate seems practical, including monetization of web content, rate limitation, and private Ethereum networks. Though several distributed web-based cryptominers have appeared in the wild (for other currencies), either in malware or in commercial trials, we believe that this is the first open-source cryptominer of this type.
% The process of legitimating cryptocurrencies (such as Bitcoin and Ethereum) is done
% through mining, often in a competitive and distributed manner by means of a small number of
% large, loosely-organized mining colonies following proprietary protocols.  We

@@

\keywords{crypto-currency \and Ethereum \and distributed computing \and web-browser computing \and mining}
\end{abstract}
%
%
%
\section{Introduction}
Cryptocurrencies are increasingly gaining traction as a viable form of currency.  This has been accompanied by a correspondingly increasing interest in the efficient validation (or \emph{mining}) of cryptocurrency transactions.  Whereas initial efforts in this domain have focused on creating dedicated hardware for this task, %AT - citation needed
\trishita{Add citation}
more recent approaches have examined repurposing existing infrastructure. Indeed, one such class of efforts has focused on the use of client web browsers as a platform for distributed computing~\cite{Cushing}. The growing popularity of CoinHive~\cite{coinhive} is a case in point of the potential success of distributed in-browser cryptocurrency mining as a commercial (if malicious) enterprise. In this work, we propose WebEth, a browser-based distributed miner for the Ethereum block chain. Obtaining a profitable hash rate within a distributed ensemble of browsers under constrained memory, compute and network usage is non-trivial. This is primarily because every worker needs to store a 2GB data structure in memory in order to mine Ethereum. However, the naive approach of transferring the data structure over the network and storing it in memory every time a browser loads our web page is not viable. And so, WebEth employs an amortized approach to mining in order to operate under such constraints. In our experiments, this approach takes at most 5 minutes to reach a steady state hash rate -- making it ideal for websites where users spend more time, like video streaming websites. WebEth yields at most 40 kiloHashes/s, which despite the overhead from running the algorithm in a browser, is only 30\% lesser than the performance of the corresponding native miner that employs our algorithm, thereby proving that our approach is effective.

\subsection{Our Contributions}
In this work, we explore the web-based cryptocurrency mining of the popular Ethereum block chain, \trishita{Add citation}
%AT - citation needed
and propose a number of potential applications consequent to this platform.  As such, our main contributions are as follows:
\begin{itemize}
\item We propose WebEth, an open-source implementation of a distributed web-based Ethereum cryptominer in both \verb|JavaScript| and \verb|WebAssembly| that can operate under relatively resource constrained environments. Though miners for other currencies exist in the wild (e.g. CoinHive~\cite{coinhive}), they are all proprietary and closed-source.
\item We propose a number of potential applications built upon such a platform, including rate-limiting server access, usage tracking, and content monetization.
\item We provide analysis and experimental evidence of the potential efficacy of our approach to the applications above. \trishita{I feel like this next part is repetitive}Our miner employs an amortized approach to mining that takes $\approx$ 5 minutes for it to reach its maximum hash rate. In addition, we show that our \verb|WebAssembly| implementation can provide hash rates up to 40 kiloHashes/s, which is only 30\% smaller than the hash rate provided by a native C++ miner that employs the same algorithm. 
\end{itemize}

@@

\subsection{Roadmap}
\trishita{Fix this acc to the new orderigns}
The remainder of the paper is organized as follows. In Section~\ref{sec:background} we cover the relevant background and related literature for our work.  In particular, we provide a self-contained description of Ethash, the Ethereum mining Proof-of-Work (PoW), and \verb|WebAssembly|, the language in which we implement an efficient miner. Section~\ref{sec:approach} outlines WebEth's approach to distributing the Ethash PoW over numerous, resource-constrained browsers. Then, we present the results in Section~\ref{sec:results}. Finally, in Section~\ref{sec:applications} we discuss various potential applications our mining platform. We conclude and discuss the future work in Sectionn~\ref{sec:conclusion}.

\section{Background}
\label{sec:background}
\subsection{Cryptocurrencies: a general overview}
% \trishita{Not sure if I need this paragraph, since I think the readers will probably be familiar with what crypto-curriencies are}

Most cryptocurrencies like Ethereum involve storing transactions in blocks, and the entire history of transactions is collated in a data structure known as a block chain. The block chain is managed by dedicated machines called ``client nodes''; where each client node typically stores the entire block chain. However, because each client node operates on its own copy of the block chain, it is very easy for the block chains on different nodes to go out of sync. Hence, there needs to be an accepted mechanism to decide the order in which new transaction blocks are appended to this block chain. 
For this, every time a new transaction takes place, the transaction is thrown in with all other transactions that have ``taken place'', but haven't been added to the block chain yet. For most currencies, the data structure that stores this is known as the ``mem pool''. A miner then picks a few transactions from this mem pool and creates a new block from these transactions. Once this happens, the goal of the miner is to have his/her block appended to the block chain, which is achieved through a process called ``mining''.
% This block chain is operated on by the Ethereum Virtual Machine (EVM) ~\cite{EVM}, which, unlike Bitcoin Script, is Turing Complete ~\cite{turingComplete}. 
% Furthermore, this currency is also a Proof-of-Work based currency, which means that the new blocks are appended to the block-chain through a process called ``mining''. 
Mining involves a race amongst miners to solve a ``Proof-of-Work puzzle'' (usually an energy intensive computation), the winner of which gets to have his block appended to the block-chain. The winner also receives a payout, which acts as an incentive to mine. 
% This process of mining is essential to deciding the order in which transactions are appended to the inherently decentralized block-chain, thereby stabilizing the chain. 
Ethereum uses Ethash as its Proof-of-Work (PoW) algorithm, which is explained in detail below.

\subsection{Ethereum and its PoW, Ethash}
Ethereum is a crypto-currency that was released in July 2015 by Vitalik Buterin, Gavin Wood and Jeffrey Wilcke. Ethereum uses the Ethash algorithm (derived from the Dagger and Hashimoto algorithms ~\cite{dagger-hashimoto}) for its PoW for mining blocks. Before we discuss how mining works with Ethash, we would first like to establish some basic terminology about the data structures involved in the PoW:
\begin{itemize}
\item \textbf{Block Header}: A block header contains meta-data related to the transactions of the block it pertains to, and is provided as an input to the Ethash algorithm.
\item \textbf{Nonce}: An integer that is provided as an input to the hashing algorithm in the hopes that when hashed with the block header, will yield a hash value less than a certain predefined difficulty threshold. It is also provided as an input to Ethash.
\item \textbf{Difficulty Threshold}: An integer value below which the hash digest output of Ethash should be in order for the miner to have successfully mined the block.
\item \textbf{Mining overview}: The Ethash mining algorithm primarily involves computing a hash of the current block header and nonce. If the output of the hash (the digest) lies below the difficulty threshold; the block is said to have been successfully mined. Hence, the miner keeps iterating over nonces until he/she gets a digest that lies below the difficulty threshold. Once the miner finds the right nonce, he/she then broadcasts the block, the header and the nonce so that the other client nodes can update their block chain to include this newly mined block. 
\item \textbf{Verification}: Once a block has been mined by a miner, it is propagated to other client nodes so that they can update their copies of the block chain. However, before each client node does so, it must validate whether the miner is submitting a legitimate block -- i.e., check whether the miner genuinely solved the hash as claimed. This is easily done by putting the block header and nonce associated with the block through the Ethash algorithm and checking whether the output indeed lies below the difficulty threshold. 
\item \textbf{Light Weight Client Nodes}
Some client nodes do not mine new blocks, but just verify whether any new block submitted by a miner is valid or not. These are called light weight client nodes.
\item \textbf{Full Client Nodes}
Full client nodes mine new blocks.
\item \textbf{Epoch}: A unit of ``time'' that spans 30,000 blocks.
\item \textbf{Seed}: All of the data structures in Ethash (mentioned below) need a 256 bit seed. This seed is different for every Epoch. For the first epoch it is the Keccak-256~\trishita{CITE} hash of a series of 32 bytes of zeros. For every other epoch it is always the Keccak-256 hash of the previous seed hash. 
\item \textbf{Cache}: The seed for a particular epoch is used to compute a data structure known as the Cache. This Cache is simply an array 4 byte integers ~\cite{Ethmining}. The cache production process involves using the seed hash to first sequentially fill up the Cache, then performing two passes of the RandMemoHash algorithm~\cite{randmemohash} on top of it to get the final Cache value. 
% This transforms the initial cache to give us the final value for the cache. 
Light weight clients may use this cache for verifying hashes (discussed below), while full node clients can use it to calculate a Directed Acyclic Graph (DAG) dataset, as described below. 
\item \textbf{DAG}: A Directed Acyclic Graph stored as a large byte array (currently around 2GB in size). The DAG has the following two attributes:
  \begin{enumerate}
  \item \textbf{Node}: Each DAG node in this byte array spans 64 bytes, and node indices are therefore aligned at a 64 byte boundary.
  \item \textbf{Page}: Each DAG page spans 2 nodes, however, page accesses are not aligned at a 2 node boundary. The mining process involves accessing some DAG pages and hashing them together with the block header and nonce.
  \end{enumerate}
Each node in the DAG is generated by combining data from 256 pseudo-randomly selected cache nodes and hashing them together. This process is repeated to generate all nodes in the DAG.
\item \textbf{Mix:} An intermediate byte array used to store temproary results in the Ethash algorithm.
\end{itemize}
We must point out that as time goes on, mining Ethereum becomes more and more difficult, as the size of the Cache and DAG increases with every Epoch.

\begin{figure}[h]
\centering
\includegraphics[width=300px,keepaspectratio]{Ethash.pdf}
\caption{\label{fig:ethash} Ethash}
\end{figure}

Mining is performed by starting with the current block header hash and nonce, which are combined to get a 128B wide ``Mix'' (another byte array),as seen in step 1 of Figure~\ref{fig:ethash}. The Mix is then used to fetch a specific page of the DAG from memory. After this, the Mix is updated with the fetched part of the DAG (step 2). Then, this updated Mix is used to fetch a new part of the DAG (step 3). This process of sequentially fetching parts of the DAG is repeated 64 times (step 4), and the final value of the Mix is put through a transformation to obtain a 32 byte digest (step 5). This digest is then compared to the threshold (step 6), and if it is smaller than it, the nonce is valid and the block is successfully mined and can be broadcast to the network. However, if the digest is greater than the threshold, the nonce is unsuccessful, and the entire process must be repeated with a new nonce ~\cite{Ethmining}.

@@

It is important to note that the parts of the DAG that are used to compute the hash for a particular block depend on the nonce used, hence there is no way to pre-determine which parts will be useful to have in memory. This therefore forces miners to store entire DAG in memory, making mining ``Memory Hard''.

@@

While mining is memory intensive, verification is relatively lightweight. This is because of the property that each node in the DAG depends on a set of pseudo-randomly selected items from the Cache. Hence, the Cache is used to regenerate only the specific nodes of the DAG that are needed to recalculate the hash for the particular nonce. And so only the Cache needs to be stored by light weight clients that only perform verification. We use this property of being able to generate parts of the DAG as needed to our advantage in order to alleviate some of the memory and network bandwidth restrictions that browsers typically face.

% Currently the DAG is more than 2GB in size, causing many GPU miners with cards of 2GB of memory to suffer in terms of hashrate. 
% The Proof-of-Work difficulty and total hashrate of the network has been steadily increasing over time, with the current difficulty at 1612.737TH and the respective hashrate at 109.138TH/s ~\cite{etherscan}. This reflects the design of mining one block approximately every 12 seconds. 
% As for mining rewards, the reward for mining a block has gone from 5 Ether to 3 ether, though the difficulty has significantly decreased, both due to the Byzantium fork.

% For a normal Ethereum miner to function, the current block headers and the DAG are required to even begin mining. As we simply cannot afford to transfer the entire DAG to the browser and store it in memory, we propose an an alternative approach to address some of these concerns.

\subsection{JavaScript and WebAssembly}
Introduced in 1995 by Netscape Communications Corporation~\cite{https://auth0.com/blog/a-brief-history-of-javascript/}, \verb|JavaScript| was a meant to be a light scripting language in order to  make web content dynamic. Over the span of 23 years, it has grown to become one of the most popular client-side web development languages used to make dynamic user interfaces.

In fact, up to recently, \verb|JavaScript| has been the only language available to make dynamic client side web content. However, the situation has changed since the advent of \verb|WebAssembly| in 2016. As per its creators, \verb|WebAssembly| is a ``binary instruction format for a stack-based virtual machine''. WASM is designed to be compiled from high-level languages like \verb|C/C++/Rust|, and is supported by 4 major browser versions -- Firefox, Safari, IE, and Chrome ~\cite{webAssembly}. The \verb|WebAssembly| stack machine is designed to be encoded in a ``size-and load-time-efficient binary format''~\cite{webAssembly}, and aims to execute near native speed by utilizing common hardware capabilities available on a wide range of platforms~\cite{webAssembly}. The language is meant to improve performance for computationally intensive use cases such as image/video editing, games, streaming, etc~\cite{webAssembly}. This, and given the fact that CoinHive is also implemented in \verb|WebAssembly| ~\cite{coinhive}, makes the language a natural choice for us to implement a miner in. 
% In fact, as of now, \verb|WebAssembly| is supported by 4 major browser versions -- Firefox, Safari, IE, and Chrome ~\cite{webAssembly}, which makes our miner implementation extremely versatile.

\subsection{Challenges}
% \trishita{Too long of a sentence; need to reword}
The crux of having a good hashrate for Ethereum lies in having quick access to the DAG. One of the biggest challenges with mining on a distributed ensemble of browsers is to transfer and store this data structure within each worker browser. However, given that the DAG for an Ethereum network is at least 1GB (and growing), transferring it to the browser every time a client loads our web page would be impractical. Furthermore, even if one somehow manages to transfer the DAG to the client when the web page loads, the browser still has to store the DAG in memory, thereby using a sizeable amount of resources on the client machine. In section~\ref{sec:approach} we discuss approaches to tackle some of these challenges.

>>>>

% remove pseudocode.
% Have a section of ethereum and define what block header and hash. Also define target etc 
% dEFINE epoch early on. 
% called Cache in ethereum terminology. Mention that the size of the cache varies right 
% define lightweight and full node earlier.
% Say each DAG consists of nodes and pages. Explain what 
% Ads: talk about how our approach can be applied to

@@

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{float}
\usepackage{graphicx}
% \usepackage{balance}
% \usepackage{color}
\usepackage{makecell}
\usepackage{amsmath}
% \usepackage{algorithmic}
\usepackage{amssymb}
\usepackage{caption}
\newcommand{\trishita}[1]{}%{{\color{magenta}\bfseries[Trishita: #1]}}
\newenvironment{claimproof}[1]{\par\noindent\underline{Proof:}\space#1}{\hfill $\blacksquare$}

% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}
\begin{document}
%
\title{Distributed Web Mining of Ethereum
% \thanks{Supported by organization x.}
}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{
Trishita Tiwari%\orcidID{0000-1111-2222-3333} 
\and
David Starobinski%\orcidID{1111-2222-3333-4444} 
\and
Ari Trachtenberg%\orcidID{2222--3333-4444-5555}
}
%
\authorrunning{T. Tiwari et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Boston University, Department of Electrical and Computer Engineering, Boston, MA 02215, USA \\
\email{\{trtiwari,staro,trachten\}@bu.edu}}
%
\maketitle             
%
\begin{abstract}
We consider the problem of mining crytocurrencies by harnessing the inherent distribution capabilities of the World Wide Web. More specifically, we propose, analyze, and implement \verb|WebEth|, a browser-based distributed miner of the Ethereum cryptocurrency. \verb|WebEth| handles Proof-of-Work (PoW) calculations through individualized code that runs on the client browsers, and thereafter collates them at a web server to complete the mining operation. \verb|WebEth| is based on an lazy evaluation technique designed to function within the expected limitations of the clients, including bounds on memory, computation and communication bandwidth to the server. We provide proofs-of-concept of \verb|WebEth| based on \verb|JavaScript| and \verb|WebAssembly| implementations, with the latter reaching hash rates up to roughly 40 kiloHashes per second, which is only 30\% slower than the corresponding native \verb|C++|-based miner.
Finally, we explore several applications of \verb|WebEth|, including monetization of web content, rate limitation to server access, and private Ethereum networks. Though several distributed web-based cryptominers have appeared in the wild (for other currencies), either in malware or in commercial trials, we believe that \verb|WebEth| is the first open-source cryptominer of this type.
% The process of legitimating cryptocurrencies (such as Bitcoin and Ethereum) is done
% through mining, often in a competitive and distributed manner by means of a small number of
% large, loosely-organized mining colonies following proprietary protocols.  We

@@

\keywords{crypto-currency \and Ethereum \and distributed computing \and web-browser computing \and mining}
\end{abstract}
%
%
%
\section{Introduction}
Cryptocurrencies are increasingly gaining traction as a viable form of currency.  This has been accompanied by a correspondingly increasing interest in the efficient validation (or \emph{mining}) of cryptocurrency transactions.  Whereas initial efforts in this domain have focused on creating dedicated hardware for this task ~\cite{asics}, %AT - citation needed
\trishita{Add citation}
more recent approaches have examined repurposing existing infrastructure. Indeed, one such class of efforts has focused on the use of client web browsers as a platform for distributed computing~\cite{Cushing}. The growing popularity of CoinHive~\cite{coinhive} is a case in point of the potential success of distributed in-browser cryptocurrency mining as a commercial (if malicious) enterprise.

In this work, we propose \verb|WebEth|, a browser-based distributed miner for the popular Ethereum block chain~\cite{ethereum}. \verb|WebEth| tackles the challenge of achieving a profitable hash rate within a distributed ensemble of browsers under constrained memory, computation and network usage. Indeed, every browser needs to store a data structure of at least 1~GB in memory in order to mine Ethereum. Clearly, it is unfeasible to transfer this entire data structure every time a browser loads a web page. Instead, \verb|WebEth| employs an lazy approach to generating this data structure while mining. Through our experiments, we show that this approach takes at most five minutes to reach a steady state hash rate -- making it ideal for web applications where users spend time, such as gaming and video streaming. Our experiments also show that \verb|WebEth| yields a hashing rate of up to 40 kiloHashes/s, which, despite the overhead from running the algorithm in a browser, is only 30\% smaller than the performance of a corresponding miner running natively.

%\subsection{Our Contributions}
%In this work, we explore the web-based cryptocurrency mining of the popular Ethereum block chain, % %\trishita{Add citation}%AT - citation needed
%and propose a number of potential applications consequent to this platform.

The main contributions of this work are as follows:
\begin{itemize}
\item We propose \verb|WebEth| \cite{WebEth}, an open-source implementation of a distributed web-based Ethereum cryptominer in both \verb|JavaScript| and \verb|WebAssembly| that can operate under relatively resource constrained environments. Though miners for other currencies exist in the wild (e.g. CoinHive~\cite{coinhive}), they are all proprietary and closed-source.
\item We provide theoretical analysis and experimental evidence of the potential efficacy of the lazy approach adopted by \verb|WebEth| to achieve high hashing rates.
%\trishita{I feel like this next part is repetitive}Our miner employs an lazy approach to mining that takes $\approx$ 5 minutes for it to reach its maximum hash rate. In addition, we show that our \verb|WebAssembly| implementation can provide hash rates up to 40 kiloHashes/s, which is only 30\% smaller than the hash rate provided by a native C++ miner that employs the same algorithm.
\item We propose a number of potential applications built upon \verb|WebEth|, including rate-limiting server access, usage tracking, and content monetization.
\end{itemize}

@@

\paragraph{Related Work}
Distribution of a common task is hardly a new concept~\cite{scheduling,parallel,orca}, but the growing popularity and efficiency of dynamic web content and client-side scripting languages like \verb|JavaScript| and \verb|WebAssembly|
have made web browsers an enticing implementation option~\cite{WebFlow,Duda,Cushing}.  Coinhive~\cite{coinhive} has
built into this environment a proprietary method for mining the Monero cryptocurrency, but this is often done on
the browsers of unsuspecting users.  A more ethical, open-sourced alternative, the Basic Attention Token~\cite{basicAttentionToken}, is an Ethereum-based ERC20 token~\cite{erc20} currently in development to be used in conjunction with the Brave browser~\cite{BraveBrowser} to generate ad revenue for website publishers by \emph{measuring a user's attention} on an advertisement. This platform promises to balance the ties between website users, publishers, and advertisers to ensure that users get only ads they would accept, advertisers pay for actual users (instead of click bots), and publishers actually get revenue instead of begging users to turn off their ad blockers.

% So far, there has been quite a lot of work done on distributed systems ~\cite{scheduling,parallel,orca}. Additionally, with the growing popularity of dynamic web-content and client side scripting languages like \verb|JavaScript| and \verb|WebAssembly|, web-browsers have also become a lucrative option for implementing distributed systems. As such, there has been some work done in exploring this relatively new avenue ~\cite{WebFlow,Duda,Cushing}. 
% Indeed, ``Distributed Computing on an Ensemble of Browsers'' by Cushing~\cite{Cushing} shows that the \verb|JavaScript| Engine has had enough improvements in performance to make computationally heavy programs feasible in a distributed web browsing unit. 
% In addition, the work by Duda and Dłubacz~\cite{Duda} was instrumental in the field as one of the first web browsing computational systems to not require the host machine to install any additional applications in order to participate. This is important since we would want our web miner to be operable without requiring the user to install any unfamiliar software. 
% Furthermore, \verb|WebAssembly|  ~\cite{webAssembly} is one of the newest platforms intended for CPU intensive browser computing (such as image rendering), and claims to be only 20\% slower than native code ~\cite{wasmPerf}, as compared to the much slower \verb|JavsScript|. Instead of relying on \verb|JavaScript|, web developers can now write C/C++ and compile it into \verb|WebAssembly| using some of the openly available \verb|WebAssembly| compilers ~\cite{emscripten}. Indeed, given that now \verb|WebAssembly| is supported by all major browsers ~\cite{webAssembly}, industry is also making use of this powerful framework.

% One sees the growing popularity of distributed web computation through the recent development of increasingly popular Monero mining library CoinHive~\cite{coinhive}. In terms of the potential monetization aspect of the project, there exists an ethical alternative to CoinHive called the Basic Attention Token ~\cite{basicAttentionToken}. The Basic Attention Token is an ERC20 Token built on Ethereum that is used along with the Brave browser to generate Ad revenue for website publishers by measuring a user's attention span on an advertisement. Note that this is not a replacement for Ad revenue, but a platform to balance the ties between website users, publishers, and advertisers. It ensures that users get only ads they would accept, advertisers pay for actual users instead of click bots, and publishers actually get revenue instead of pleading the users to turn off their Ad blockers.

\paragraph{Roadmap}
The remainder of the paper is organized as follows. In Section 2\ref{sec:background} we cover the relevant background and related literature for our work.  In particular, we provide a self-contained description of Ethash, the Ethereum mining Proof-of-Work (PoW), and \verb|WebAssembly|, the language in which we implement an efficient miner. Section 3\ref{sec:webeth} describes \verb|WebEth|, including its lazy approach to distributing the Ethash PoW over numerous, resource-constrained browsers. This section also includes a performance analysis. We present the experimental results of our implementations in Section 4\ref{sec:results}. In Section 5\ref{sec:applications} we discuss several potential applications of our mining platform. We conclude in Section 6\ref{sec:conclusion}.

\section{Background}
\label{sec:background}
\subsection{Cryptocurrencies: a general overview}
Most cryptocurrencies like Ethereum involve storing transactions in blocks, and the entire history of transactions is collated in a data structure known as a block chain. The block chain is managed by dedicated machines called \emph{client nodes}; where each client node typically stores the entire block chain. However, because each client node operates on its own copy of the block chain, the block chains on different nodes may go out of sync. Hence, there needs to be an accepted mechanism to decide the order in which new transaction blocks are appended to this block chain. 
For this, every time a new transaction takes place, the transaction is pooled together with all other transactions that have been broadcasted to the network, but haven't been added to the block chain yet. For most currencies, the data structure that stores these unconfirmed transactions is known as the \emph{mem pool}. A miner picks valid transactions from his mem pool and creates a new block out of these transactions. Once this happens, the goal of the miner is to have his/her block appended to the block chain, which is achieved through a process called \emph{mining}.
% This block chain is operated on by the Ethereum Virtual Machine (EVM) ~\cite{EVM}, which, unlike Bitcoin Script, is Turing Complete ~\cite{turingComplete}. 
% Furthermore, this currency is also a Proof-of-Work based currency, which means that the new blocks are appended to the block-chain through a process called ``mining''. 
Mining involves a race amongst miners to solve a \emph{Proof-of-Work} (PoW) puzzle, which is usually an energy intensive computation. The winner of this race gets to have his block appended to the block chain. The winner also receives a payout, which acts as an incentive to mine. 
% This process of mining is essential to deciding the order in which transactions are appended to the inherently decentralized block-chain, thereby stabilizing the chain. 
Ethereum uses Ethash as its PoW algorithm, which is explained in detail below.

\subsection{Ethereum Proof of Work}
Ethereum is a crypto-currency that was released in July 2015 by Vitalik Buterin, Gavin Wood and Jeffrey Wilcke. Ethereum uses the Ethash algorithm (derived from the Dagger and Hashimoto algorithms ~\cite{dagger-hashimoto}) for its PoW for mining blocks. Before we discuss how mining works with Ethash, we first establish basic terminology about the data structures involved in the PoW.

A \emph{block header} contains meta-data related to the transactions of the corresponding block, and it is provided as an input to the Ethash algorithm together with an integer \emph{nonce}.  The nonce is chosen in a brute-force fashion in order to hash, together with the block header, into a value that matches a specific pattern (based on a predefined \emph{difficulty threshold}). The process of finding an appropriate nonce for a given block is known as \emph{mining}.
%\item \textbf{Difficulty Threshold}: An integer value below which the hash digest output of Ethash should be in order for the miner to have successfully mined the block.

Once a block has been mined, it is propagated to other client nodes so that they can update their copies of the block chain. However, before each client node does so, it must validate whether the miner is submitting a legitimate block -- i.e., check whether the miner genuinely solved the hash as claimed. This is easily done by putting the block header and nonce associated with the block through the Ethash algorithm and checking whether the output follows the pattern prescribed by the difficulty threshold.  \emph{Light weight} client nodes do not mine new blocks, but rather only verify whether any new block submitted by a miner is valid or not.  \emph{Full} client nodes, on the other hand, both mine and verify new blocks.

An \emph{epoch} is a unit of ``time'' that spans 30,000 blocks.  All of the data structures in Ethash (mentioned below) need a 256 bit \emph{seed}, which differs for every epoch. For the first epoch the seed is a Keccak-256~\cite{keccak} hash of a series of 32 bytes of zeros. For every other epoch it is always the Keccak-256 hash of the previous seed hash.  The seed for a particular epoch is used to compute a data structure known as the \emph{Cache}, which is an array of 4 byte integers \cite{Ethmining}. The Cache production process involves using the seed hash to first sequentially fill up the Cache, then performing two passes of the RandMemoHash algorithm~\cite{randmemohash} on top of it to get the final Cache value. 
% This transforms the initial cache to give us the final value for the cache. 
Light weight clients may use this Cache for verifying hashes, while full node clients can use it to calculate a Directed Acyclic Graph (DAG) dataset, as described below.  A Ethereum \emph{Directed Acyclic Graph} is stored as a large byte array (around 1 GB in size on the private Ethereum Network used for our experiments) and has the following two attributes:
  \begin{enumerate}
  \item \emph{Node}: Each DAG node in this byte array spans 64 bytes, and node indices are therefore aligned at a 64 byte boundary.
  \item \emph{Page}: Each DAG page spans 2 nodes, however, page accesses are not aligned at a 2 node boundary. The mining process involves accessing some DAG pages and hashing them together with the block header and nonce.
  \end{enumerate}
Each node in the DAG is generated by combining data from 256 pseudo-randomly selected Cache nodes and hashing them together. This process is repeated to generate all nodes in the DAG. Finally, intermediate byte arrays used to store temporary results in the Ethash algorithm are known as \emph{Mix}es.

We must point out that as time goes on, mining Ethereum becomes more and more difficult, as the size of the Cache and DAG increases with every epoch.

\begin{figure}[h]
\centering
\includegraphics[width=300px,keepaspectratio]{Ethash.pdf}
\caption{Ethash}
\label{fig:ethash}
\end{figure}

Mining is performed by starting with the current block header hash and nonce, which are combined to get a 128-byte wide Mix, as seen in step 1 of Figure 1\ref{fig:ethash}. The Mix is then used to fetch a specific page of the DAG from memory. After this, the Mix is updated with the fetched page of the DAG (step 2). Then, this updated Mix is used to fetch a new page of the DAG (step 3). This process of sequentially fetching parts of the DAG is repeated 64 times (step 4), and the final value of the Mix is put through a transformation to obtain a 32 byte digest (step 5). This digest is then compared to the threshold (step 6). If it is smaller than the threshold, the nonce is valid and the block is successfully mined and can be broadcast to the network. However, if the digest is greater than the threshold, the nonce is unsuccessful, and the entire process must be repeated with a new nonce~\cite{Ethmining}.

@@

It is important to note that the pages of the DAG that are used to compute the hash for a particular block depend on the nonce used, hence there is no way to pre-determine which pages will be useful to have in memory. This therefore forces miners to store entire DAG in memory, making Ethereum mining ``Memory Hard''.

@@

Whereas mining is memory intensive, verification is relatively lightweight. This is because of the property that each node in the DAG depends on a set of pseudo-randomly selected items from the Cache. Hence, the Cache is used to regenerate only the specific pages of the DAG that are needed to recalculate the hash for the particular nonce. And so, only the Cache needs to be stored by light weight clients that only perform verification. In fact, as we will see in the next section, we use this property of being able to generate parts of the DAG as needed to our advantage in order to alleviate some of the memory and network bandwidth restrictions that browsers typically face.

\subsection{JavaScript and WebAssembly}
Introduced in 1995 by Netscape Communications Corporation~\cite{js}, \verb|JavaScript| was a meant to be a light scripting language in order to  make web content dynamic. Over the span of 23 years, it has grown to become one of the most popular client-side web development languages used to make dynamic user interfaces.

In fact, up to recently, \verb|JavaScript| has been the only language available to make dynamic client-side web content. However, the situation has changed since the advent of \verb|WebAssembly| in 2016. As per its creators, \verb|WebAssembly| is a ``binary instruction format for a stack-based virtual machine'' \cite{webAssembly}. WASM is designed to be compiled from high-level languages like \verb|C/C++/Rust|, and is supported by 4 major browser platforms -- Firefox, Safari, IE, and Chrome \cite{webAssembly}. The \verb|WebAssembly| stack machine is designed to be encoded in a ``size-and load-time-efficient binary format''~\cite{webAssembly}, and aims to execute near native speed by utilizing common hardware capabilities present on a wide variety of platforms~\cite{webAssembly}. The language is meant to improve performance for computationally intensive use cases such as image/video editing, games, streaming, etc~\cite{webAssembly}. This makes it the language of choice to implement a miner within a browser.

%\subsection{Challenges}
%The crux of having a good hashrate for Ethereum lies in having quick access to the DAG. One of the biggest challenges with mining on a distributed ensemble of browsers is to transfer and store this data structure within each worker browser. However, given that the DAG for an Ethereum network is at least 2 GB (and growing), transferring it to the browser every time a client loads our web page is impractical. Furthermore, even if one somehow manages to transfer the DAG to the client when the web page loads, the browser still has to store the DAG in memory, thereby using a sizeable amount of resources on the client machine. In section~\ref{sec:approach} we discuss approaches to tackle some of these challenges.

\section{WebEth}
\label{sec:webeth}
%\label{sec:approach}
%\subsection{Approach}
% \trishita{I think it would be nice if we show how we modified the algorithm shown in the previous section. I will add that over here}

\verb|WebEth| is based on a lazy evaluation to alleviate the network and memory requirements for mining Ethereum in a distributed scenario on browsers. Specifically, as soon as each browser connects to the webserver, the server sends to the browser the current header of the block being mined and the Cache. Once the browser receives the Cache, it allocates an array buffer to store the DAG node.
%s that fall between indices 0 and $N$ that will be computed in the future. (Note that the $N$ is a variable parameter reflecting the length of the buffer with which we experiment in our analysis). 
Once the buffer is allocated, the browser can start iterating over nonces to compute hashes.

Since, to begin with, the browser does not store any nodes of the DAG, it must compute each node on the fly using the Cache. However, every node that the browser computes is stored in the buffer, for quicker access in the future. Hence, as time passes, the buffer starts filling up, such that more and more nodes are quickly accessed from the buffer rather than being computed from the ground-up, which makes hash computations faster with time. This has the effect that the longer the user remains on the web-page, the better the hash rate gets for that user.
